---
title: "Writing to a DB"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Writing to a DB}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`faketables` offers `dbWriteTable()` as a convenient way to write modified data back to a data source.

In the below example, `dbConnect()` and `dbDisconnect()` should produce a [DBI Connection Object](https://dbi.r-dbi.org/reference/dbConnect.html) and disconnect from that database, respectively. The database connection and disconnection is not required to be within the observe statement; the only requirement is that the connection passed to `dbWriteTable` be an active `DBIConnection` object.

```{r, eval=FALSE}
f_tab <- faketables::faketable()

ui <- shiny::fluidPage(
  shiny::actionButton('db_write', 'Write to DB')
)

server <- function(input, output, session) {
  f_tab <- faketables::faketablesServer(f_tab)
  
  shiny::observe({
    con <- dbConnect()
    faketables::dbWriteTable(con, 'table_name', f_tab)
    dbDisconnect(con)
  }) |>
    shiny::bindEvent(input$db_write)
}
```

A more complete example using the same [Favorite Pizza Places app](https://landeranalytics.github.io/faketables/articles/favorite_pizza.html) can be found [here](https://raw.githubusercontent.com/landeranalytics/faketables/main/app_db.R).

