---
title: "faketables"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{faketables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Intro

This vignette is a complete example and shows how to use `faketables` and all of its components.

In this vignette we will create a shiny app that allows us to modify data from our dataset and then preview each component of our `faketables` object.

There is a more in-depth example available [here](https://landeranalytics.github.io/faketables/articles/favorite_pizza.html).

## Setup

Load our library.

```{r}
library(faketables)
```

To make sure things don't get out of hand, we first filter down to just the first six rows. Then, we want the rownames as a column, and then for the sake of example, we convert `qsec` to a list column where every row is a vector of length two that is the original value of `qsec` both subtracted by and added by `1`. We then select the columns we wish to display.

```{r}
df <-
  mtcars |>
  head() |>
  tibble::rownames_to_column() |>
  dplyr::mutate('qsec' = purrr::map(.data$qsec, \(x) round(c(x - 1, x + 1)))) |>
  dplyr::select('rowname', 'mpg', 'cyl', 'qsec')
```

## Creating a `faketable`

### Understanding `input_call()`

Because `faketables` creates the inputs dynamically, it can't work with the output of a function, but rather needs to work with the function itself. This is called a "bare function". `input_call()` requires a bare function and a list of `args` to be passed to the bare function.

The bare function usually takes the form of the function call, but without the parenthesis or any of the arguments such as `sum` for the `sum()` function. Alternatively, a function definition can be passed. It is important that this match exactly, or the `updated` table will contain all rows.

The `args` should be a list of all desired arguments that would normally be passed to your function, except the `inputId`. Even if one is supplied, it will be ignored.

These can be stored as a variable or passed directly to the `input` argument of `col_def()`.

### Creating a `col_def`

First, we need a `col_def` for each column we wish to display. We will name ours `c_def`. Each `col_def` must have the `name` of the column it is being created for, an `input_call` object, a bare function call to `cast` and ensure that the column has the desired class, a `width`. They can also have an optional `display_name`.

There are some things to note here:

- Our label is `NULL` so that the input isn't labelled for each row, but it doesn't have to be that way.
- `'rowname'` uses an anonymous function that wraps the input it `shinyjs::disabled()`. This is so the `rowname` column can be rendered as part of the table without users being allowed to modify its value.
- `'qsec'` uses an anonymous function that wraps the actual casting function. This is because `qsec` should be a list column, but we still want to ensure the contents of the list match our desired class.

```{r}
c_def <- list(
  col_def(
    name = 'rowname',
    input = input_call(
      fun = \(inputId, ...) { shinyjs::disabled(shiny::textInput(inputId, ...)) },
      args = list(label = NULL, placeholder = 'Vehicle Name')
    ),
    cast = as.character,
    width = 3,
    display_name = 'Vehicle Name'
  ),
  col_def(
    name = 'mpg',
    input = input_call(
      fun = shiny::numericInput,
      args = list(label = NULL)
    ),
    cast = as.numeric,
    width = 2,
    display_name = 'MPG'
  ),
  col_def(
    name = 'cyl',
    input = input_call(
      fun = shiny::selectInput,
      args = list(label = NULL, choices = c(4, 6, 8))
    ),
    cast = as.integer,
    width = 2,
    display_name = 'CYL'
  ),
  col_def(
    name = 'qsec',
    input = input_call(
      fun = shiny::sliderInput,
      args = list(label = NULL, min = 10, max = 25)
    ),
    cast = \(x) purrr::map(x, as.numeric),
    width = 3,
    display_name = 'QSEC'
  )
)
```

### Creating a `table_def`

A `table_def` is, at its core, a collection of `col_def` objects that have been neatly organized into a `tibble`. We will name ours `t_def`.

```{r}
t_def <- table_def(c_def)
```

### Creating a `faketable`

The last step in creating our `faketable` is to bring our data and `table_def` together. Since we know our `rowname` column is unique, we use that as our `rowId`. We also supply an empty list so that the delete buttons are show for each row.

```{r}
f_tab <- faketable(df, t_def, rowId = 'rowname', show_delete = list())
```

## The Shiny App

### Creating the UI

Our UI will be relatively simple as we're just trying to demonstrate our `faketable`. First, we use `faketablesUI()` to make sure our `faketable` is rendered. Then we create a row of inputs that match those from our `col_def` so that we can input data and press the `'Add Row'` button to add the row. At the bottom, we have a radio button that lets us switch between each of our `faketable` object tables to preview.

```{r}
ui <- shiny::fluidPage(
  title = 'mtcars',
  shinyjs::useShinyjs(),
  faketablesUI(),
  shiny::tags$h3('Add Row'),
  shiny::fluidRow(
    shiny::column(
      width = 2, shiny::actionButton('add_row', 'Add Row')
    ),
    shiny::column(
      width = 3, shiny::textInput('rowname', 'Vehicle Name')
    ),
    shiny::column(
      width = 2, shiny::textInput('mpg', 'MPG', value = 50)
    ),
    shiny::column(
      width = 2, shiny::selectInput('cyl', 'CYL', c(4, 6, 8))
    ),
    shiny::column(
      width = 3, shiny::sliderInput('qsec', 'QSEC', min = 10, max = 25, value = c(10, 25))
    )
  ),
  shiny::tags$h3('Preview Table'),
  shiny::radioButtons(
    'preview', 
    label = 'Choose Table to Preview', 
    choices = c('data', 'inserted', 'updated', 'deleted'),
    inline = TRUE
  ),
  shiny::tableOutput('preview_table')
)
```

### Creating the Server

The first thing we do is call `faketablesServer()` on our `faketable`, `f_tab`, and assign it back to `f_tab`. This is no longer a `faketables` object, but a [shiny::reactive()](https://shiny.posit.co/r/reference/shiny/latest/reactive.html) that holds our `faketables` object. As such, in order to access our `faketable`, we now have to use `f_tab()` rather than `f_tab`. This reactive will handle all row updating and deleting on its own. In order to support inserting rows, we need to create an event listener for the `'Add Row'` button and call our `faketablesInsert` using our now reactive `f_tab` and the new data we want to insert.


```{r}
server <- function(input, output, session) {
  f_tab <- faketablesServer(faketable = f_tab)
  
  # get the table referenced by the radio buttons
  # `object@property_name` is the same as `prop(object, 'property_name')`
  preview_table <- shiny::reactive({ S7::prop(f_tab(), input$preview) })
  # use an un-exported function to ensure list columns render
  output$preview_table <- shiny::renderTable(faketables:::.list_col_to_chr(preview_table()))
  
  shiny::observe({
    ins <- tibble::tibble(
      'rowname' = input$rowname,
      'mpg' = as.numeric(input$mpg),
      'cyl' = as.numeric(input$cyl),
      'qsec' = list(as.integer(input$qsec))
    )
    faketables::faketablesInsert(f_tab, ins)
  }) |>
    shiny::bindEvent(input$add_row)
}
```

### Running the app

```{r eval=FALSE}
shiny::shinyApp(ui, server)
```

