---
title: "Inserting Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Inserting Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`faketables` provides two methods for inserting data. If you're trying to insert data in a `Shiny` app, you want `faketablesInsert()`. Otherwise, you may want `insert()`.

## `insert()`

`insert()` is for use on a `faketables` object and does the actual data insertion. This is called automatically in `faketablesServer()` when the `insert` argument is supplied data. Alternatively, if `faketables` is being used outside of a `Shiny` context for its data history properties, it can be used to safely add new data to the `faketables` object.

Usage is simply `insert(faketable, new_data)`.

## `faketablesInsert()`

`faketablesInsert()` is for use in `Shiny` apps when the app developer wants new data added to the `faketables` object represented by a `reactive` object. Just like `insert()`, it is provided two arguments. The latter of which is the new data to be inserted. However, unlike `insert()`, the first argument is not a `faketables` object, but rather the variable that represents the output of a `faketablesServer()` call.

Usage is a little bit more complex than that of `insert()` and requires two components. The first is that the server call and assign `faketablesServer()` and the second is a call to `faketablesInsert` where the new data is being actively added to the `faketables` object.

```{r, eval=FALSE}
f_tab <- faketables::faketable()
server <- function(input, output, session) {
  f_tab <- faketables::faketablesServer(f_tab)
  
  shiny::observe({
    faketables::faketablesInsert(f_tab, new_data)
  }) |>
    shiny::bindEvent(input$new_data_button)
}
```

